---
title: "HNET analysis"
author: "Philipp Chapkovski"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r set - options, echo = FALSE, cache = FALSE}
options(width = 5000)
options(digits = 5)
options(scipen = 50)
options(dplyr.summarise.inform = F)
knitr::opts_chunk$set(
  echo = FALSE,
  cache = T,
  warning = F,
  results = 'markup',
  message = FALSE
)

```

``` {r libraries}
library('pacman')
p_load('dplyr', 'estimatr', 'tidyverse', 'readxl', 'sjPlot', 'emmeans', 'olsrr', 'writexl', 'gtsummary', 'stargazer', 'magrittr', 'ggplot2', 'scales','lemon','knitr','ggpmisc', 'fastDummies', 'estimatr', 'sjPlot','haven',
       'hrbrthemes','kableExtra', 'skimr','lubridate', 'data.table', 'gganimate','transformr', 'png','gifski','summarytools', 'dm', 'DBI', 'glue')

```

```{r dbconnect}


con <- dbConnect(odbc::odbc(),
                 driver = "/usr/local/lib/libsqlite3odbc.dylib",
                 database = "/Users/chapkovski/archive/hnet_scraper/db.sqlite3")
dm <- dm_from_src(con)

dm<-dm %>% 
  dm_add_pk(hnet_structuredpost,id) %>% 
  dm_add_pk(hnet_position,id) %>% 
  dm_add_fk(hnet_structuredpost_positions,structuredpost_id, hnet_structuredpost)%>%
  dm_add_fk(hnet_structuredpost_positions,position_id, hnet_position)%>%
  dm_add_pk(hnet_category,id) %>% 
  dm_add_fk(hnet_structuredpost_primary_categories,structuredpost_id, hnet_structuredpost)%>%
  dm_add_fk(hnet_structuredpost_primary_categories,category_id, hnet_category)%>%
    dm_add_fk(hnet_structuredpost_secondary_categories,structuredpost_id, hnet_structuredpost)%>%
  dm_add_fk(hnet_structuredpost_secondary_categories,category_id, hnet_category)




drop.cols=c('created_at', 'updated_at' )
position_tab <- dm %>%
  dm_select(hnet_position, -one_of(drop.cols)) %>%
  dm_select(hnet_structuredpost, -one_of(drop.cols))%>%
  dm_disambiguate_cols(sep='.')%>%
  dm_flatten_to_tbl(hnet_structuredpost_positions) %>%
  as_tibble()%>%
  rename(position=description)%>%
  mutate(position=str_trim(position)) %>%
    mutate(year=lubridate::year(posting_date),
         month=lubridate::month(posting_date),
         day=lubridate::day(posting_date),
         weekday=lubridate::wday(posting_date),
         week=lubridate::isoweek((posting_date)),
         monthdate = as.Date(paste("2020", month, '01', sep = "-")),
         fixeddate = as.Date(paste("2020", month, day, sep = "-"))
         )

# Preparing primary category analysis
# 

primary_cat_tab<-dm %>%
  dm_select(hnet_category, -one_of(drop.cols)) %>%
  dm_select(hnet_structuredpost, -one_of(drop.cols))%>%
  dm_disambiguate_cols(sep='.')%>%
  dm_flatten_to_tbl(hnet_structuredpost_primary_categories) %>%
  as_tibble()%>%
  rename(category=description)%>%
  mutate(category=str_trim(category)) %>%
    mutate(year=lubridate::year(posting_date),
         month=lubridate::month(posting_date),
         day=lubridate::day(posting_date),
         weekday=lubridate::wday(posting_date),
         week=lubridate::isoweek((posting_date)),
         monthdate = as.Date(paste("2020", month, '01', sep = "-")),
         fixeddate = as.Date(paste("2020", month, day, sep = "-"))
         )

# Preparing secondary category analysis
# 

secondary_cat_tab<-
  dm %>%
  dm_select(hnet_category, -one_of(drop.cols)) %>%
  dm_select(hnet_structuredpost, -one_of(drop.cols))%>%
  dm_disambiguate_cols(sep='.')%>%
  dm_flatten_to_tbl(hnet_structuredpost_secondary_categories) %>%
  as_tibble() %>%
  rename(category=description)%>%
  mutate(category=str_trim(category)) %>%
    mutate(year=lubridate::year(posting_date),
         month=lubridate::month(posting_date),
         day=lubridate::day(posting_date),
         weekday=lubridate::wday(posting_date),
         week=lubridate::isoweek((posting_date)),
         monthdate = as.Date(paste("2020", month, '01', sep = "-")),
         fixeddate = as.Date(paste("2020", month, day, sep = "-"))
         )



```

## Number of posts per year

```{r preparing data for posts}
posts<-as_tibble(tbl(con, "hnet_structuredpost") )%>%
  mutate(year=lubridate::year(posting_date),
         month=lubridate::month(posting_date),
         day=lubridate::day(posting_date),
         weekday=lubridate::wday(posting_date),
         week=lubridate::isoweek((posting_date)),
         monthdate = as.Date(paste("2020", month, '01', sep = "-")),
         fixeddate = as.Date(paste("2020", month, day, sep = "-"))
         ) %>%
  group_by(week,year)%>%
  mutate(weekdate=min(posting_date),
         fixedweek=ymd(format(weekdate, "2020-%m-%d")))%>%
  ungroup()
```

```{r plotting }




posts %>%
  mutate(monthdate=ymd(glue('{year}-{month}-01')),
         yeardate=ymd(glue('{year}-01-01')))%>%
  group_by(yeardate)%>%
  summarise(posts=n())%>%
  ggplot(aes(x=yeardate, y=posts, fill=yeardate))+
  geom_col(width = 200.8, position = position_dodge(width = 200.9)  )+
  geom_label(aes(label=posts, y=posts/2 ),colour='white', size=3)+
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
   theme(axis.text.x = element_text(angle = 90))+xlab('')


```
Posts per year with moving average for 3 months

## Posts per week per year 

```{r polarplot}



averages<-posts %>%
  filter(year<2020)%>%
  group_by(week, year) %>% 
  summarise(posts=n()) %>%
  group_by(week)%>%
  summarise(averages=mean(posts))


byweek<-posts %>%
  group_by(year,fixedweek,week)%>%
  summarise(posts= n())
  full_join(byweek,averages) %>%
  mutate(col=case_when(year==2020~'red',TRUE~'lightgray'))%>%
  ggplot(aes(x=fixedweek, y=posts, group=year)) +
  ylim(0,NA) +
  geom_line(aes(color=col))+
  geom_line(data=full_join(byweek,averages)%>%filter(year==2020),aes(x=fixedweek, y=averages, group=NA),color='blue')+
  coord_polar(start =0) +
  scale_color_identity() +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  xlab('Week of the year')+
  ylab('Job ads posted per week')+
  theme_ipsum()+
  labs(title = "Posts per week")

```

## Seasonality (posts per month)

```{r posts-per-month}


posts %>%
  group_by(monthdate, year) %>% summarise(n = n()) %>% group_by(monthdate) %>%
  mutate(lab = case_when(n == max(n) | n == min(n) ~ year),
         vj= case_when(n == max(n) ~ -1.5 , n == min(n) ~ 1.5)) %>%
  ggplot(aes(x = monthdate, y = n)) +
  geom_point(, stat = 'identity') +
  geom_smooth(method = "loess") +
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
  xlab('Month') +
  ylab('Number of posts') + geom_text(aes(label = lab, vjust = vj))+
  ylim(0,500)


```
## Dynamics of professorship positions vs other positions

In the plot below percentages is a share of professorship positions in total number of job ads posted per year.

For the last 10 years share of professorship positions are in decline. It substantially fell in 2020 to less than a half. We might believe it's because other positions are posted more often, but in fact total number of posts per year is also in decline (we of course cannot guarantee that it's not due to falling popularity of H-Net website as an advertising platform) 

```{r positions}

position_tab %>%
  mutate(prof=str_detect(position, 'Professor')) %>%
  mutate(prof = if_else(prof,  "Yes", "No")) %>%
  group_by(prof, year)%>%
  tally() %>%
  filter(year>2003)%>%
  group_by(year)%>%
  mutate(freq=if_else(prof=='Yes', n/sum(n), NA_real_))%>%
  ggplot(aes(x=year, y=n, fill=prof))+
  geom_col()+
  geom_label(aes(label=label_percent(accuracy=1)(freq)))+
  labs(fill='Professor positions')+
  theme_ipsum()
  


```

## Black studies/Race studies positions

```{r cat-black}

black_obj<-list(filter_str='frican|Race|lack',
                title='Black/African/African-American', col='navyblue')


graph_builder <- function(data, obj, ordertype, rel=F) {
  if (rel){
    reltitle <- 'Share'
    custom_scale<-  scale_y_continuous(labels = label_percent(accuracy=1))

  } else {
    reltitle <- 'Number'
    custom_scale<-list()
  }
  
  subchain <- . %>% 
    group_by(year)%>%
  mutate(n=n/sum(n))
  
  data %>%
    mutate(black = str_detect(category, obj$filter_str)) %>%
    group_by(black,year) %>%
    tally() %>%
   {if(rel) subchain(.) else .} %>%
    filter(black == T) %>%
    ggplot(aes(x = year,  y = n)) +
    geom_segment(aes(
      x = year,
      xend = year,
      y = 0,
      yend = n
    ), color = "lightgreen") +
    geom_point(
      aes(x = year, y = n),
      fill = obj$col,
      colour = 'black',
      size = 4 ,
      pch = 21
    ) +
    custom_scale +
    ggtitle(glue("{reltitle}  of {obj$title}  positions"), subtitle=glue('({ordertype})'))
}

graph_builder(primary_cat_tab, black_obj, 'primary', rel=F)
graph_builder(secondary_cat_tab, black_obj, 'secondary', rel=F)

graph_builder(primary_cat_tab, black_obj, 'primary', rel=T)
graph_builder(secondary_cat_tab, black_obj, 'secondary', rel=T)



```

## Feminism/Women studies positions

```{r cat-women}

women_obj<-list(filter_str='wom|Wom|fem|Fem|Gend|gend',title='Women/Gender', col='pink')
graph_builder(primary_cat_tab, women_obj, 'primary', rel=F)
graph_builder(secondary_cat_tab, women_obj, 'secondary', rel=F)

graph_builder(primary_cat_tab, women_obj, 'primary', rel=T)
graph_builder(secondary_cat_tab, women_obj, 'secondary', rel=T)



```



## Positions posted by continent and country

```{r locations}

# RUN THE FOLLOWING JUST ONCE BECAUSE IT REQUESTS A BUNCH OF THINGS FROM GEOCODING API (*TWICE*); THE RESULTS ARE SAVED IN .csv 
# 
# p_load('ggmap','revgeo')
# 
# api_key<-'YOUR API KEY'
# register_google(key = api_key)


# loc2country <- function(loc) {
#   latlon <-
#     ggmap::geocode(location = c(loc),
#                    output = "latlon",
#                    source = "google") %>% as_tibble()
#   revgeo::revgeo(
#     longitude = latlon$lon,
#     latitude = latlon$lat,
#     provider = 'google',
#     API = api_key,
#     output = 'hash',
#     item = 'country'
#   )$country
# }

# alllocs<-alllocs %>%
#   arrange(location)%>%
#   mutate(country=loc2country(location))

# alllocs$continent <- countrycode(sourcevar = alllocs$ country,
#                             origin = "country.name",
#                             destination = "continent")

# write.csv(alllocs, 'alllocs.csv')
alllocs<- read.csv('alllocs.csv')
alllocs <- alllocs %>% mutate(country=recode(country, `Yugoslavia`='Kosovo'))
left_join(posts, alllocs) %>%
  filter(year%in%c(2007,2020))%>%
  group_by(continent, year)%>%
  tally() %>%
  group_by(year)%>%
  mutate(freq=n/sum(n))%>%
  pivot_wider(names_from=year,
              values_from=freq,
              id_cols=continent) %>%
   mutate(mean = rowMeans(across(where(is.numeric))))%>%
  arrange(mean)%>%
  mutate(x=factor(continent),
         value1=`2007`,
         value2=`2020`)%>%
  ggplot()+
  geom_segment( aes(x=x, xend=x, y=value1, yend=value2), color="grey") +
  geom_point( aes(x=x, y=value1), color=rgb(0.2,0.7,0.1,0.5), size=3 ) +
  geom_point( aes(x=x, y=value2), color=rgb(0.7,0.2,0.1,0.5), size=3 ) +
  coord_flip()+
  theme_ipsum() +
  theme(
    legend.position = "none",
  ) +
  xlab("") +
  ylab("Value of Y")+
  geom_text(aes(x=x,   y=value1, vjust=2, label=label_percent(accuracy=1)(value1)))+
  geom_text(aes(x=x,   y=value2, vjust=2, label=label_percent(accuracy=1)(value2)))+
    geom_text(aes(x=x,   y=value1, vjust=-2, label='2007'))+
  geom_text(aes(x=x,   y=value2, vjust=-2, label='2020'))+
  ylim(0,1.2)+
  scale_y_continuous(labels = label_percent())

```


```{r positions-absolute}
```
